---
title: "NYC Schools Perceptions"
author: "Mykola Zotko"
data: "January 20, 2019"
output: html_notebook

---

In this project we will work with data sets from the [New York City Department of Education](https://www.schools.nyc.gov/). They contain data on NYC schools, including student demographics, test scores, graduation and dropout rates, and school locations.
We will incorporate some additional data from responses to surveys designed to gauge parent, student, and teacher perceptions of the quality of New York City schools. 

```{r, message=FALSE}
library(dplyr)
library(readr)

```
Read preproceseed NYC schools data, [servey data](https://data.cityofnewyork.us/Education/2011-NYC-School-Survey/mnz3-dyi8) for all community schools and for all District 75 schools.

```{r, message=FALSE}
combined <-  read_csv('data/combined.csv')
survey <-  read_tsv('data/masterfile11_gened_final.txt')
survey_d75 <- read_tsv('data/masterfile11_d75_final.txt')

```

Filter `survey` data to include only high schools and select columns needed for analysis.

```{r}
survey <- survey %>%
  filter(schooltype == "High School") %>%
  select(dbn, schoolname, rr_s:p_q3b)

```
Select columns needed for analysis from `survey_d75`.

```{r}
survey_d75 <- survey_d75 %>%
  subset(is.na(highschool)) %>%
  select(dbn, rr_s:p_q3c)
```

